Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-01-23 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex

The upper bound $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$ on the expected number of
mistakes of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} is
optimal up to a constant factor, as long as the number of classes $K$ is at most
$O((R/\gamma)^2)$. This lower bound is stated as
\autoref{theorem:strongly-separable-examples-mistake-lower-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.
\citet{Daniely-Helbertal-2013} provide a lower bound under the assumption of weak linear separability, which does not immediately imply a lower bound under the stronger notion.

@@

\paragraph{Remark.}
If $\gamma \le R$ then, irrespective of any other conditions on $K$, $R$, and
$\gamma$, a trivial lower bound on the expected number of mistakes of any
randomized algorithm is $(K-1)/2$. To see this, note that the adversary can
choose an example $(R e_1, y)$, where $e_1$ is some arbitrary unit vector in $V$
and $y$ is a label chosen uniformly from $\cbr{1,2,\dots,K}$, and show this
example $K$ times. The sequence of examples trivially satisfies the strong
linear separability condition, and the $(K-1)/2$ expected mistake lower bound
follows from ~\citep[][Claim 2]{Daniely-Helbertal-2013}. %However, it is unclear
%if the trivial lower bound is the best possible if $K$ is large, i.e.,
%$\omega((R/\gamma)^2)$. We leave this as an open problem.

>>>>

The upper bound $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$ on the expected number of
mistakes of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} is
optimal up to a constant factor, as long as the number of classes $K$ is at most
$O((R/\gamma)^2)$. This lower bound is stated as
\autoref{theorem:strongly-separable-examples-mistake-lower-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

@@

\paragraph{Remark.}
\citet{Daniely-Helbertal-2013} provide a lower bound under the assumption of weak linear separability, which does not immediately imply a lower bound under the stronger notion.
If $\gamma \le R$ then, irrespective of any other conditions on $K$, $R$, and
$\gamma$, a trivial lower bound on the expected number of mistakes of any
randomized algorithm is $(K-1)/2$. To see this, note that the adversary can
choose an example $(R e_1, y)$, where $e_1$ is some arbitrary unit vector in $V$
and $y$ is a label chosen uniformly from $\cbr{1,2,\dots,K}$, and show this
example $K$ times. The sequence of examples trivially satisfies the strong
linear separability condition, and the $(K-1)/2$ expected mistake lower bound
follows from ~\citep[][Claim 2]{Daniely-Helbertal-2013}. %However, it is unclear
%if the trivial lower bound is the best possible if $K$ is large, i.e.,
%$\omega((R/\gamma)^2)$. We leave this as an open problem.

### 2019-01-23 ICML-submission/related-work.tex

Many works consider bandit multiclass classification in the general setting,
where the examples are not necessarily separable by some linear
classifier~\cite{Abernethy-Rakhlin-2009, Wang-Jin-Valizadegan-2010,
Crammer-Gentile-2013, Hazan-Kale-2011, Beygelzimer-Orabona-Zhang-2017,
Foster-Kale-Luo-Mohri-Sridharan-2018}. In addition, \citet{Chen-Lin-Lu-2014,
Zhang-Jung-Tewari-2018} study online bandit multiclass boosting under bandit
feedback, where one can view boosting as linear classification by treating each
base hypothesis as a separate feature. However, most of these works achieve a
regret of order $\widetilde{O}(\sqrt{T})$ to $O(T^{3/4})$. In the weakly
linearly separable setting, these algorithms at best guarantees a mistake upper
bound of $\widetilde{O}(\sqrt{T})$, which is undesirable for problems with a
long time horizon. One exception is the \textsc{Newtron} algorithm, where with
appropriate settings of parameter $\alpha$, it is able to achieve a $O(\ln T)$
regret against the $\alpha$-logistic loss. However, as argued in
~\citep[][Appendix D]{Beygelzimer-Orabona-Zhang-2017}, if the setting of
$\alpha$ makes the regret bound $O(\sqrt{T})$, the $\alpha$-logistic loss of the
best linear predictor would be $\Omega(T^{3/4})$. Therefore, \textsc{Newtron} at
best guarantees a mistake upper bound of $O(\sqrt{T})$ in the weakly linearly
separable setting.

@@

There is a large body of literature on the contextual bandits
problem~\citep{Auer-2003, Langford-Zhang-2008}. In this problem, the learner is
given a policy class $\Pi$, and at each timestep $t$, an example $x_t$ is shown,
with an associated cost vector $c_t$ hidden. The learner then takes an action
$\widehat{y}_t$, and observes the incurred cost $c_t(\widehat{y}_t)$. The goal
of the learner is to minimize its regret $\sum_{t=1}^T c_t(\widehat{y}_t) -
\min_{\pi \in \Pi} \sum_{t=1}^T c_t(\pi(x_t))$. The bandit multiclass learning
problem can be seen as a special case, where the cost $c_t(i)$ is the
classification error $\indicator{i \neq y_t}$, and the policy class $\Pi$ is the
set of linear classifiers $\cbr{x \mapsto \argmax_y (Wx)_y: W \in \R^{k \times
d}}$. A series of works in the contextual bandits learning literature assume
access to a policy optimization oracle that returns a policy in $\Pi$ that
minimizes its empirical cost on any set of cost-sensitive examples, and
construct algorithms that call the oracle a polynomial number of times%
~\citep{Dudik-Hsu-Kale-Karampatziakis-Langford-Reyzin-Zhang-2011,
Agarwal-Hsu-Kale-Langford-Li-Schapire-2014, Rakhlin-Sridharan-2016,
Syrgkanis-Krishnamurthy-Schapire-2016,
Syrgkanis-Luo-Krishnamurthy-Schapire-2016}. However, these algorithms are not
truly computationally efficient in our setting, as it is NP-hard in general to
find a linear classifier that minimizes the empirical cost over a set of
examples~\citep{Arora-Babai-Stern-Sweedyk-1997}.

>>>>

The general problem of linear bandit multiclass learning has received considerable attention~\cite{Abernethy-Rakhlin-2009, Wang-Jin-Valizadegan-2010,
Crammer-Gentile-2013, Hazan-Kale-2011, Beygelzimer-Orabona-Zhang-2017,
Foster-Kale-Luo-Mohri-Sridharan-2018}. \citet{Chen-Lin-Lu-2014,
Zhang-Jung-Tewari-2018} study online bandit multiclass boosting under bandit
feedback, where one can view boosting as linear classification by treating each
base hypothesis as a separate feature.  
%However, most of these works achieve a
%regret of order $\widetilde{O}(\sqrt{T})$ to $O(T^{3/4})$.  
In the weakly linearly separable setting, however, 
all these algorithms guarantee a mistake 
bound of ${O}(\sqrt{T})$ at best.
%One exception is the \textsc{Newtron} algorithm, which---with
%an appropriate setting of parameter $\alpha$---can achieve a $O(\ln T)$
%regret against the $\alpha$-logistic loss. However, as argued in
%~\citep[][Appendix D]{Beygelzimer-Orabona-Zhang-2017}, if the setting of
%$\alpha$ makes the regret bound $O(\sqrt{T})$, the $\alpha$-logistic loss of the
%best linear predictor would be $\Omega(T^{3/4})$. Therefore, \textsc{Newtron} at
%best guarantees a mistake upper bound of $O(\sqrt{T})$ in the weakly linearly
%separable setting.

@@

There is a large body of literature on contextual bandit
learning~\citep{Auer-2003, Langford-Zhang-2008}. In this problem, the learner is
given a policy class $\Pi$, and at each timestep $t$, an example $x_t$ is shown,
with an associated cost vector $c_t$ hidden. The learner then takes an action
$\widehat{y}_t$, and observes the incurred cost $c_t(\widehat{y}_t)$. The goal
of the learner is to minimize its regret $\sum_{t=1}^T c_t(\widehat{y}_t) -
\min_{\pi \in \Pi} \sum_{t=1}^T c_t(\pi(x_t))$. The bandit multiclass learning
problem can be seen as a special case, where the cost $c_t(i)$ is the
classification error $\indicator{i \neq y_t}$, and the policy class $\Pi$ is the
set of linear classifiers $\cbr{x \mapsto \argmax_y (Wx)_y: W \in \R^{k \times
d}}$. A series of papers in the contextual bandits learning literature assume
access to a policy optimization oracle that returns a policy in $\Pi$ that
minimizes its empirical cost on any set of cost-sensitive examples, and
construct algorithms that call the oracle a polynomial number of times%
~\citep{Dudik-Hsu-Kale-Karampatziakis-Langford-Reyzin-Zhang-2011,
Agarwal-Hsu-Kale-Langford-Li-Schapire-2014, Rakhlin-Sridharan-2016,
Syrgkanis-Krishnamurthy-Schapire-2016,
Syrgkanis-Luo-Krishnamurthy-Schapire-2016}. However, these algorithms are not
truly computationally efficient in our setting, as it is NP-hard in general to
find a linear classifier that minimizes the empirical cost over a set of
examples~\citep{Arora-Babai-Stern-Sweedyk-1997}.
